---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 150
  )
```
# nycgeo

[![Travis build status](https://travis-ci.org/mfherman/nycgeo.svg?branch=master)](https://travis-ci.org/mfherman/nycgeo)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/nycgeo)](https://cran.r-project.org/package=nycgeo)
  
The `nycgeo` package contains spatial data files for various geographic and administrative boundaries in New York City. Data is in the [`sf` (simple features)](https://r-spatial.github.io/sf/) format and includes boundaries for boroughs (counties), public use microdata areas (PUMAs), neighborhood tabulation areas (NTAs), and census tracts. In the future, more boundaries will be added, such as community districts, census blocks, and city council districts.

In addition, selected demographic, social, and economic estimates from the U.S. Census Bureau American Community Survey can be added to the geographic boundaries in `nycgeo`, allowing for contextualization and easy choropleth mapping. Finally, `nycgeo` makes it simple to access a subset of spatial data in a particular geographic area, such as all census tracts in Brooklyn and Queens.

## Why `nycgeo`?

The spatial files contained in the `nycgeo` package are available on websites such as the [New York City Department of City Planning's Bytes of the Big Apple](https://www1.nyc.gov/site/planning/data-maps/open-data.page#district_political) and the [U.S. Census Bureau TIGER/Line® Shapefiles](https://www.census.gov/geo/maps-data/data/tiger-line.html), but this package aims to make accessing the spatial data more convenient. Instead of downloading and converting shapefiles each time you need them, `nycgeo` provides the files in a consistent format (`sf`) with added metadata that enable joins with non-spatial data.

Other R packages share some features with `nycgeo`. In particular, the wonderful [`tidycensus`](https://walkerke.github.io/tidycensus/) package can access the Census Bureau's API and download ACS estimates as well as TIGER/Line® Shapefiles (via [`tigris`](https://github.com/walkerke/tigris)). 

One difference between the boundaries included here and the TIGER/Line® Shapefiles available through `tigris` is that these boundaries are clipped to the shoreline, allowing for better mapping of New York City. Additionally, `nycgeo` contains boundaries for geographic areas that are not available from the Census Bureau. This includes neighborhood tabulation areas (NTAs) and community districts (CDs).

Finally, all spatial data included in the package uses the [NAD83 / New York Long Island (ftUS) State Plane projected coordinate system (EPSG 2263)](https://epsg.io/2263), which is the standard projection used by New York City government agencies.


## Installation

You can install `nycgeo` from [GitHub](https://https://github.com/mfherman/nycgeo) with:

``` r
remotes::install_github("mfherman/nycgeo")
```

## Examples

### Basic Usage

```{r load-packages, include = FALSE, echo = FALSE}
library(nycgeo)
library(sf)
library(tidyverse)
```

The most basic usage of `nycgeo` is to get boundaries in the `sf` format. Use `nyc_tracts()`, `nyc_pumas()`, or `nyc_ntas()` to get your desired geography. To make best use of the package, you should also load the `sf` package when using `nycgeo`. For these examples, I'll also load `tidyverse` as this will allow us to take advantage of pretty `tibble` printing and will come in handy when we want to manipulate and map the spatial data later.

```{r basic-usage}
library(nycgeo)
library(sf)
library(tidyverse)

nyc_tracts()
```

### Filter by geography

If you don't need census tracts for the entire city, you can use the `filter_by` and `region` arguments of `nyc_tracts()` to specify the area you are interested in. For example, the following code returns only census tracts in Brooklyn and Queens.

```{r filter-by}
bk_qn_tracts <- nyc_tracts(
  filter_by = "borough",
  region = c("brooklyn", "queens")
  )

ggplot(bk_qn_tracts) +
  geom_sf() +
  theme_minimal()
```

Note, you can select multiple regions by passing a character vector to the `region` argument, but you can only select a single geography to `filter_by`. Additionally, you can only filter by a geography that is *larger than or equal to* the boundaries you request. For example, it is not possible to filter PUMAs by NTAs because NTAs are smaller than PUMAs.

### Adding American Community Survey Data

`nycgeo` includes selected estimates from the American Community Survey as datasets. You can access these datasets directly or have them appended to the spatial data. To print a `tibble` of ACS data, simply call the data you want.

```{r print-acs-data}
ntas_acs_data
```

To add census estimates to an `sf` object, use `add_acs_data = TRUE` in the appropriate `nyc_*()`call. For example, here we get all NTAs in Manhattan with ACS data appended. One convenience of having the ACS data joined to the `sf` object is that you can very simply make a choropleth map. Here we do it with `ggplot2`, but you could use `tmap`, `leaflet` or any other spatial package that works with `sf` objects.

```{r add-acs-data}
mn_ntas <- nyc_ntas(
  filter_by = "borough",
  region = "manhattan",
  add_acs_data = TRUE
  )

ggplot(mn_ntas) +
  geom_sf(aes(fill = pop_ba_above_pct_est)) +
  scale_fill_viridis_c(
    name = "Bachelor's or above",
    labels = scales::percent_format(),
    option = "magma"
    ) +
  theme_void() +
  theme(panel.grid = element_line(color = "transparent")) +
  labs(title = "Which neighborhoods in Manhattan are most educated?")
```

### Joining with other data

One use case of `nycgeo()` is if you have non-spatial data that relates to census tracts, NTAs, or other geographies and need to join that data with spatial boundaries to plot or otherwise analyze. This non-spatial data may be coded in a variety of ways and might not have names or IDs that match your spatial data. The `sf` data provided in `nycgeo` seeks to have a variety of geographic metadata that will match whatever labels your non-spatial data has. 

In this example, we have non-spatial data from the [NYC Neighborhood Health Atlas](https://www1.nyc.gov/site/doh/health/neighborhood-health/nyc-neighborhood-health-atlas.page) at the NTA-level from which we would like to make a choropleth map. To do this, we import the .csv file and then join it to the spatial NTA object matching on NTA IDs. Then, we can map it as in the above example.

```{r join-health, message = FALSE}
nta_health <- read_csv("https://raw.githubusercontent.com/mfherman/nycgeo/master/inst/extdata/nta-health.csv") %>% 
  select(NTA_Code, BlackCarbon)

nyc_ntas() %>% 
  left_join(nta_health, by = c("nta_id" = "NTA_Code")) %>% 
  ggplot() +
  geom_sf(aes(fill = BlackCarbon)) +
  scale_fill_viridis_c(name = "Black carbon (absorbance units)", option = "inferno") +
  theme_void() +
  theme(panel.grid = element_line(color = "transparent")) +
  labs(title = "Which neighborhoods have high levels of black carbon pollution?")
```